---
title: "Scripps's Murrelet Egg Size Model"
author: "Amelia J. DuVall & Marcela Todd Zaragoza"
date: "2021"
output: pdf_document
---

This is `r paste0("v.", (Sys.Date()))`

# Introduction
This document includes code for likelihood ratio tests to test inclusion of random effects, model selection, and model diagnostics for a linear-mixed model of Scripps's Murrelet *Synthliboramphus scrippsi* egg size as a function of laying sequence and oceanographic indices at Santa Barbara Island, Channel Islands National Park, USA from 2009-2017.  
\
```{r setup, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load libraries
library(here)
library(tidyverse)
library(janitor)
library(ggplot2)
library(lubridate)
library(viridis)
library(lme4)
library(RLRsim)
library(faraway)
library(sjPlot)
library(forecast)

## load egg size data and covariate data 
egg <- read.csv(here("data", "SCMU_egg_data.csv"), sep = ";") %>%
  mutate(Year = as.numeric(Year)) # make year an integer
covars <- read.csv(here("data", "covariates", "covars.csv"))

## join covariate data with egg size data 
SCMUdf <- left_join(egg, covars, by = "Year") %>% # join by year
  filter(TrueOrder == TRUE) %>% # select eggs with laying sequence known only
  filter(Size > 1210) %>% # remove the outliers
  dplyr::select(Year, Observer, Plot, Size, LayingSequence, ANCHL, BEUTI, NPGO, ONI, PDO, SST) 
```

# Global Model
```{r mods}
## create data frame specifying predictors to include
predictors <- as.data.frame(matrix(c(FALSE, TRUE), 2, 7)) 
# 7 potential predictors (includes LayingSequence)

## add column names
cov_names <- colnames(predictors) <- colnames(SCMUdf[,5:11])

## create set of all possible combinations
full_set <- expand.grid(predictors) 

## select models with correlated predictors
ii <- which(full_set$ANCHL + full_set$NPGO == 2 |
              full_set$ANCHL + full_set$ONI == 2 |
              full_set$ANCHL + full_set$PDO == 2 |
              full_set$BEUTI + full_set$ONI == 2 |
              full_set$BEUTI + full_set$PDO == 2 |
              full_set$NPGO + full_set$PDO == 2 |
              full_set$NPGO + full_set$SST == 2 |
              full_set$ONI + full_set$PDO == 2) 

## create reduced set of models and convert to a matrix for easier indexing
use_set <- as.matrix(full_set[-ii,]) 

## number of models in set
(n_mods <- nrow(use_set)) 

## find max number of predictors in a model
max(rowSums(use_set))

## covariates in global model
cov_names[use_set[which.max(rowSums(use_set)),]]

```

# Likelihood Ratio Tests
We used likelihood ratio tests using the [RLRsim](https://cran.r-project.org/web/packages/RLRsim/RLRsim.pdf) package on the global model to test the support for inclusion of two random effects: 

1) Plot: the monitoring plot from which the egg was obtained and measured (n = 8)  
2) Observer: the person who measured the egg (n = 27)  
\
```{r lrt}
## remove NAs in Observer field (e.g., unknown Observer)
SCMUdf2 <- SCMUdf[-c(which(is.na(SCMUdf$Observer==TRUE))),]

## global model with both random effects
bm_both <- lmer(Size ~ LayingSequence + ANCHL + BEUTI + SST + 
                  (1 | Observer) + (1 | Plot), data = SCMUdf2, REML = FALSE)

## run model with plot random effect only
bm_plot <- lmer(Size ~ LayingSequence + ANCHL + BEUTI + SST + 
                  (1 | Plot), data = SCMUdf2, REML = FALSE)

## run model with observer random effect only
bm_obs <- lmer(Size ~ LayingSequence + ANCHL + BEUTI + SST + 
                 (1 | Observer), data = SCMUdf2, REML = FALSE)
# Error "(boundary(singular) fit: see ?isSingular" due to very small or zero variance; does not affect model results. 

## Exact RLRT test
# m is the reduced model containing only the RE to be tested with 
# the random effect set to zero under the null hypothesis. 
# mA and Mo are the models under the alternative and the null, respectively. 

# observer set to zero under the null hypothesis
exactRLRT(m = bm_obs, mA = bm_both, m0 = bm_plot, seed = 16)

# plot set to zero under the null hypothesis 
exactRLRT(m = bm_plot, mA = bm_both, m0 = bm_obs, seed = 16)
```

# Model Diagnostics
These diagnostics are done for the global model. 
\
```{r global}
## run global model
global_mod <- lmer(Size ~ LayingSequence + ANCHL + BEUTI + SST + (1 | Plot), 
                   data = SCMUdf, REML = TRUE)
```

## Predicted Egg Size Values
```{r d1}
## extract predicted egg size values and plot
preds <- predict(global_mod)
p1 <- ggplot() + 
  geom_histogram(mapping = aes(preds), bins = 15) + # set bins 
  theme_minimal() +
  xlab("Predicted Egg Size") + ylab("Count") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
p1
```

## Residuals
```{r d2}
## extract residuals and plot
resids <- residuals(global_mod)
p2 <- ggplot() + 
  geom_histogram(mapping = aes(resids), bins = 15) +
  theme_minimal() +
  xlab("Residuals") + ylab("Count") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
p2
```

## Model Coefficients
```{r d3}
## extract coeffs and random effects
coef(global_mod) # this include fixed and random effects
ranef_pl <- ranef(global_mod)$Plot # plot random effect only

## look at data going into random effects
table(SCMUdf$Plot)
```

## Q-Q Plots
```{r d4}
qqresids <- as.data.frame(resids)
p3 <- ggplot(qqresids, aes(sample = resids)) + 
  geom_qq() + geom_qq_line() +
  theme_minimal() +
  xlab("Theoretical Quantiles") + ylab("Sample Quantiles") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("QQ Plot (Residuals)")
p3

plotRE <- as.data.frame(unlist(ranef_pl))
p4 <- ggplot(plotRE, aes(sample = unlist(ranef_pl))) + 
  geom_qq() + geom_qq_line() +
  theme_minimal() +
  xlab("Theoretical Quantiles") + ylab("Sample Quantiles") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(hjust = 0.5)) +
  ggtitle("QQ Plot (Plot Random Effect)")
p4
```

## Fitted Values Versus Residuals
```{r d5}
## plot residuals versus fitted values
yh <- fitted(global_mod)
plot(yh, resids, las = 1, pch = 16,
     xlab = "Fitted", ylab = "Residuals",
     main = "Residuals vs fitted")
abline(h=0, lty = "dashed")
```

## Levene's Test
We can formally test the assumption of homogenous variance via the Levene's Test, which compares the absolute values of the residuals among groups. 
\
```{r lev}
## split residuals into 2 groups
g1 <- resids[yh <= median(yh)]
g2 <- resids[yh > median(yh)]

## Levene's test
var.test(g1, g2)
```

# Fit Candidate Models
```{r modsel}
## create empty matrix for storing results
mod_res <- matrix(NA, n_mods, 1)
colnames(mod_res) <- c("AIC")

## fit models and store AIC 
for(i in 1:n_mods) {
  if(i == 1) {
    fmla <- "Size ~ 1 + (1 | Plot)"
  } else {
    fmla <- paste("Size ~ (1 | Plot) +", paste(cov_names[use_set[i,]], collapse = " + "))
  }
  mod_fit <- lmer(as.formula(fmla), data = SCMUdf, REML = TRUE)
  mod_res[i,"AIC"] <- AIC(mod_fit)
}

## create empty matrix for storing results
delta_res <- matrix(NA, n_mods, 1)
colnames(delta_res) <- c("deltaAIC")

## convert IC to deltaIC
delta_res[,"deltaAIC"] <- mod_res[,"AIC"] - min(mod_res[,"AIC"])
(delta_res <- round(delta_res, 2)) # round results

# top 5 models
top6 <- order(delta_res)[1:6]

## "best" models from our set
cov_names[use_set[top6[1],]] # LayingSequence, BEUTI, NPGO
cov_names[use_set[top6[2],]] # LayingSequence, NPGO, ONI
cov_names[use_set[top6[3],]] # LayingSequence, ANCHL, BEUTI, SST (>2 AIC)
cov_names[use_set[top6[4],]] # LayingSequence, PDO, SST (>2 AIC)
cov_names[use_set[top6[5],]] # LayingSequence, PDO (>2 AIC)
cov_names[use_set[top6[6],]] # LayingSequence, BEUTI, SST (>2 AIC)

## run top models
topmod1 <- lmer(Size ~ LayingSequence + BEUTI + NPGO + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
topmod2 <- lmer(Size ~ LayingSequence + NPGO + ONI + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
topmod3 <- lmer(Size ~ LayingSequence + ANCHL + BEUTI + SST + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
topmod4 <- lmer(Size ~ LayingSequence + PDO + SST + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
topmod5 <- lmer(Size ~ LayingSequence + PDO + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
topmod6 <-lmer(Size ~ LayingSequence + BEUTI + SST + (1 | Plot), 
               data = SCMUdf, REML = TRUE)

## Model selection table
AIC.tab <- matrix(NA, nrow = 6, ncol = 3) # 6 rows for 6 top models
AIC.tab[1,1] <- AIC(topmod1) # AIC for topmod1 in first row, first column
AIC.tab[2,1] <- AIC(topmod2) # AIC for topmod2 in second row, first column
AIC.tab[3,1] <- AIC(topmod3) # AIC for topmod3 in second row, first column
AIC.tab[4,1] <- AIC(topmod4) # AIC for topmod4 in second row, first column
AIC.tab[5,1] <- AIC(topmod5) # AIC for topmod5 in second row, first column
AIC.tab[6,1] <- AIC(topmod6) # AIC for topmod6 in second row, first column
AIC.tab[,2] <- AIC.tab[,1] - min(AIC.tab[,1]) # calculate delta AIC
AIC.tab[,3] <- exp(-0.5*AIC.tab[,2])/
  (sum(exp(-0.5*AIC.tab[,2]))) # calculate model weights
colnames(AIC.tab) <- c("AIC", "deltaAIC", "model_weights")
print(AIC.tab)
```

# Top Models

## Top Model 1
```{r topmod1}
topmod1 <- lmer(Size ~ LayingSequence + BEUTI + NPGO + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
summary(topmod1)  
confint(topmod1)
```

## Top Model 2
```{r topmod2}
topmod2 <- lmer(Size ~ LayingSequence + NPGO + ONI + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
summary(topmod2)
confint(topmod2)
```

## Top Model 3
```{r topmod3}
topmod3 <- lmer(Size ~ LayingSequence + ANCHL + BEUTI + SST + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
summary(topmod3)
confint(topmod3)
```

## Top Model 4
```{r topmod4}
topmod4 <- lmer(Size ~ LayingSequence + PDO + SST + (1 | Plot), 
                data = SCMUdf, REML = TRUE)
summary(topmod4)
confint(topmod4)
```