---
title: "Single Covariate Model Selection for Scripps's Murrelet Linear Mixed Model"
author: "Amelia J. DuVall & Marcela Todd"
date: "5/16/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load libraries
library(here)
library(tidyverse)
library(janitor)
library(ggplot2)
library(lubridate)
library(viridis)
library(lme4)
library(faraway)
library(sjPlot)
```

This is `r paste0("v.", (Sys.Date()))`

# Introduction
This document tests the inclusion of lagged and temporally restricted environmental covariates for the Scripps's Murrelet egg size model (see SCMU_egg_model.Rmd). We initially selected the time period January-June for covariates available at a monthly sampling rate due to the overlap with the murrelet breeding season. However, we felt it was pertinent to tests lags on all environmental covariate data since the effects of some environmental conditions may take time to manifest in terms of prey availability for murrelets during the breeding season. For each environmental covariate, we used single covariate models to test the support of a non-lagged covariate (2009-2017 data) versus a one-year lagged covariate (2008-2016 data) via AIC model selection. Most covariates were available at a monthly sampling rate, with the exception of larval anchovy. For covariates at a monthly sampling rate, we also tested a "full" covariate model with monthly sampling data averaged across 12 months versus a "half" covariate model with monthly sampling data averaged across 6 months (January-June).

We tested 6 covariates: 
1. Larval anchovy - SCC (ANCHL)  
2. Biologically Effective Upwelling Transport Index (BEUTI)  
3. Sea Surface Temperature (SST)  
4. North Pacific Gyre Oscillation Index (NPGO)  
5. Pacific Decadal Oscillation Index (PDO)  
6. Oceanic Nino Index (ONI)  

See SCMU_egg_covariates.Rmd for more information on covariate sourcing.  

```{r df}
## load egg size data
egg <- read.csv(here("data", "SCMU_egg_data.csv"))%>%
  filter(TrueOrder == TRUE) %>% # egg order known only
  select(Year, Observer, Plot, Size, EggOrder)
```

# Environmental Covariates

## ANCHL
ANCHL sampling data is only available at a yearly sampling interval. 
```{r anchovyl}
## load data
ANCHLraw  <- read.csv(here("data", "covariates", "cciea_EI_FBS_2020_a29e_1fd0_409d.csv"), na.strings = "NaN")

## are there NAs in the data?
sum(is.na(ANCHLraw$relative_abundance))

## clean data and filter down to years 2008-2017
ANCHL <- ANCHLraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(ANCHL = as.numeric(relative_abundance)) %>%
  arrange(Year) %>%
  filter(Year >= 2008 & Year <= 2017) %>%
  dplyr::select(Year, ANCHL) 

## create no lagged dataset from 2009-2017 and scale values
ANCHL_nolag <- ANCHL %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(ANCHL_nolag = scale(ANCHL)) %>%
  dplyr::select(Year, ANCHL_nolag)

## created one-year lagged dataset from 2008-2016 and scale values
ANCHL_lag <- ANCHL %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(TrueYear = Year,
         Year = TrueYear + 1,
         ANCHL_lag = scale(ANCHL)) %>%
  dplyr::select(Year, ANCHL_lag)

## join no-lag and lag datasets
ANCHL1 <- full_join(ANCHL_nolag, ANCHL_lag, by = "Year")

## join with egg size data
ANCHLdf1 <- left_join(egg, ANCHL1, by = "Year")
  
## run models
ANCHL_nolag_mod <- lmer(Size ~ ANCHL_nolag + (1 | Plot), data = ANCHLdf1) 
ANCHL_lag_mod <- lmer(Size ~ ANCHL_lag + (1 | Plot), data = ANCHLdf1)

## compare AIC
ANCHL_nolag_AIC <- AIC(ANCHL_nolag_mod)
ANCHL_lag_AIC <- AIC(ANCHL_lag_mod)

## True or False, is the ANCHL no lag model AIC smaller than the ANCHL lag AIC?
ANCHL_nolag_AIC < ANCHL_lag_AIC

## What is the delta AIC?
ANCHL_lag_AIC - ANCHL_nolag_AIC
```
There is support to use the non-lagged ANCHL data since the lagged data is delta AIC > 2. 

## BEUTI
First, we tested against using data from the entire year versus from January-June. 
```{r BEUTI1}
BEUTIraw <- read_csv(here("data", "covariates", "cciea_OC_BEUTI_784c_ef9f_af6f.csv"))

## are there NAs in the data?
sum(is.na(BEUTIraw$beuti))

## Average sampling data by year over all 12 months and scale
BEUTI_full <- BEUTIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(beuti = as.numeric(beuti)) %>%
  group_by(Year) %>%
  summarise(BEUTI = mean(beuti, na.rm = TRUE)) %>%
  dplyr::select(Year, BEUTI) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(BEUTI_full = scale(BEUTI)) %>%
  arrange(Year) %>%
  dplyr::select(Year, BEUTI_full)

## Average sampling data by year for January-June only and scale
BEUTI_half <- BEUTIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(beuti = as.numeric(beuti)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>%
  group_by(Year) %>%
  summarise(BEUTI = mean(beuti, na.rm = TRUE)) %>%
  dplyr::select(Year, BEUTI) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(BEUTI_half = scale(BEUTI)) %>%
  arrange(Year) %>%
  dplyr::select(Year, BEUTI_half)

## join 2 datasets together
BEUTI1 <- full_join(BEUTI_full, BEUTI_half, by = "Year")

## join with egg size data
BEUTIdf1 <- left_join(egg, BEUTI1, by = "Year")
  
## run models
BEUTI_full_mod <- lmer(Size ~ BEUTI_full + (1 | Plot), data = BEUTIdf1) 
BEUTI_half_mod <- lmer(Size ~ BEUTI_half + (1 | Plot), data = BEUTIdf1)

## compare AIC
BEUTI_full_AIC <- AIC(BEUTI_full_mod)
BEUTI_half_AIC <- AIC(BEUTI_half_mod)

## True or False, is the Jan-Jun BEUTI model AIC smaller than the BEUTI full AIC?
BEUTI_half_AIC < BEUTI_full_AIC

## What is the delta AIC?
BEUTI_half_AIC - BEUTI_full_AIC
```
Both models are supported since they are within 2 delta AIC. 

Then, using the half-year model, we tested a non-lagged versus one-year lagged dataset.
```{r BEUTI2}
## no lag, data averaged for Jan-June (same as BEUTI_half above) and scaled
BEUTI_nolag <- BEUTIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(beuti = as.numeric(beuti)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>%
  group_by(Year) %>%
  summarise(BEUTI = mean(beuti, na.rm = TRUE)) %>%
  dplyr::select(Year, BEUTI) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(BEUTI_nolag = scale(BEUTI)) %>%
  arrange(Year) %>%
  dplyr::select(Year, BEUTI_nolag)

## one year lag, data averaged for Jan-June and scaled
BEUTI_lag <- BEUTIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(beuti = as.numeric(beuti)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>%
  group_by(Year) %>%
  summarise(BEUTI = mean(beuti, na.rm = TRUE)) %>%
  dplyr::select(Year, BEUTI) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(BEUTI_lag = scale(BEUTI),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, BEUTI_lag)

## join 2 datasets together
BEUTI2 <- full_join(BEUTI_nolag, BEUTI_lag, by = "Year")

## join with egg size data
BEUTIdf2 <- left_join(egg, BEUTI2, by = "Year")
  
## run models
BEUTI_nolag_mod <- lmer(Size ~ BEUTI_nolag + (1 | Plot), data = BEUTIdf2) 
BEUTI_lag_mod <- lmer(Size ~ BEUTI_lag + (1 | Plot), data = BEUTIdf2)

## compare AIC
BEUTI_nolag_AIC <- AIC(BEUTI_nolag_mod)
BEUTI_lag_AIC <- AIC(BEUTI_lag_mod)

## True or False, is the no lag BEUTI model AIC smaller than the BEUTI lag AIC?
BEUTI_nolag_AIC < BEUTI_lag_AIC

## What is the delta AIC?
BEUTI_nolag_AIC - BEUTI_lag_AIC
```
There is more support for a one-year lagged BEUTI with the half-year data versus a no-lag BEUTI with the half year data. 

Next, we compared a full one-year lagged BEUTI versus half one-year lagged BEUTI.
```{r BEUTI3}
## one year lag, averaged across 12 months and scaled
BEUTI_full_lag <- BEUTIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(beuti = as.numeric(beuti)) %>%
  group_by(Year) %>%
  summarise(BEUTI = mean(beuti, na.rm = TRUE)) %>%
  dplyr::select(Year, BEUTI) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(BEUTI_full_lag = scale(BEUTI),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, BEUTI_full_lag)

## one year lag, averaged for Jan-June (same above BEUTI_lag above) and scaled
BEUTI_half_lag <- BEUTIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(beuti = as.numeric(beuti)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>%
  group_by(Year) %>%
  summarise(BEUTI = mean(beuti, na.rm = TRUE)) %>%
  dplyr::select(Year, BEUTI) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(BEUTI_half_lag = scale(BEUTI),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, BEUTI_half_lag)

## join 2 datasets together
BEUTI3 <- full_join(BEUTI_full_lag, BEUTI_half_lag, by = "Year")

## join with egg size data
BEUTIdf3 <- left_join(egg, BEUTI3, by = "Year")

## run model
BEUTI_full_lag_mod <- lmer(Size ~ BEUTI_full_lag + (1 | Plot), data = BEUTIdf3) 
BEUTI_half_lag_mod <- lmer(Size ~ BEUTI_half_lag + (1 | Plot), data = BEUTIdf3) 
  
## compare AIC
BEUTI_full_lag_AIC <- AIC(BEUTI_full_lag_mod)
BEUTI_half_lag_AIC <- AIC(BEUTI_half_lag_mod)

## True or False, is the half lag BEUTI model AIC smaller than the BEUTI full lag AIC?
BEUTI_half_lag_AIC < BEUTI_full_lag_AIC

## What is the delta AIC?
BEUTI_full_lag_AIC - BEUTI_half_lag_AIC
```
Both models are supported (which is what we would expect based on the first two tests). There is support to use 12-month averaged lagged data OR 6-month averaged lagged data for BEUTI. 

## SST
First, we tested against using data from the entire year versus from January-June. 
```{r sst1}
SSTraw <- read_csv(here("data", "covariates", "cciea_OC_SST3_91cf_d165_213f-46025.csv"))

## are there NAs in the data?
sum(is.na(SSTraw$SST))

## Averaged over all 12 months and scaled
SST_full <- SSTraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(sst = as.numeric(SST)) %>%
  group_by(Year) %>%
  summarise(SST = mean(sst, na.rm = TRUE)) %>%
  dplyr::select(Year, SST) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(SST_full = scale(SST)) %>%
  arrange(Year) %>%
  dplyr::select(Year, SST_full)

## Averaged for January-June only and scaled
SST_half <- SSTraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(sst = as.numeric(SST)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>% 
  group_by(Year) %>%
  summarise(SST = mean(sst, na.rm = TRUE)) %>%
  dplyr::select(Year, SST) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(SST_half = scale(SST)) %>%
  arrange(Year) %>%
  dplyr::select(Year, SST_half)

## join 2 datasets together
SST1 <- full_join(SST_full, SST_half, by = "Year")

## join with egg size data
SSTdf1 <- left_join(egg, SST1, by = "Year")

## run models
SST_full_mod <- lmer(Size ~ SST_full + (1 | Plot), data = SSTdf1) 
SST_half_mod <- lmer(Size ~ SST_half + (1 | Plot), data = SSTdf1)

## compare AIC
SST_full_AIC <- AIC(SST_full_mod)
SST_half_AIC <- AIC(SST_half_mod)

## True or False, is the Jan-Jun SST model AIC smaller than the SST full AIC?
SST_half_AIC < SST_full_AIC

## What is the delta AIC?
SST_half_AIC - SST_full_AIC
```
The full-year SST model has more support than the half-year model.

Then, using the top model, we tested a non-lagged versus one-year lagged dataset.
```{r SST2}
## no lag, averaged for 12 months (same as SST_full above) and scaled
SST_nolag <- SSTraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(sst = as.numeric(SST)) %>%
  group_by(Year) %>%
  summarise(SST = mean(sst, na.rm = TRUE)) %>%
  dplyr::select(Year, SST) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(SST_nolag = scale(SST)) %>%
  arrange(Year) %>%
  dplyr::select(Year, SST_nolag)

## one year lag, averaged for 12 months and scaled
SST_lag <- SSTraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(sst = as.numeric(SST)) %>%
  group_by(Year) %>%
  summarise(SST = mean(sst, na.rm = TRUE)) %>%
  dplyr::select(Year, SST) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(SST_lag = scale(SST),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, SST_lag)

## join 2 datasets together
SST2 <- full_join(SST_nolag, SST_lag, by = "Year")

## join with egg size data
SSTdf2 <- left_join(egg, SST2, by = "Year")
  
## run models
SST_nolag_mod <- lmer(Size ~ SST_nolag + (1 | Plot), data = SSTdf2) 
SST_lag_mod <- lmer(Size ~ SST_lag + (1 | Plot), data = SSTdf2)

## compare AIC
SST_nolag_AIC <- AIC(SST_nolag_mod)
SST_lag_AIC <- AIC(SST_lag_mod)

## True or False, is the no lag SST model AIC smaller than the SST lag AIC?
SST_nolag_AIC < SST_lag_AIC

## What is the delta AIC?
SST_lag_AIC - SST_nolag_AIC
```
The non-lagged SST data has more support than the one-year lagged SST data. There is support for a non-lagged SST data from the entire year (12 months). 

## NPGO
First, we tested three single covariate models: 1) data averaged from the entire year, 2) data averaged from January-June, and 3) data averaged from December-March (as supported by Chenillat et al. 2012, "the statistical relationship between NPGO index and nearshore wind variability (most upwelling favorable) along the U.S. West coast is strongest in the wintertime (December to March) off Central California"). 
```{r NPGO}
NPGOraw <- read_csv(here("data", "covariates", "cciea_OC_NPGO_712b_5843_9069.csv"))

## are there NAs in the data?
sum(is.na(NPGOraw$NPGO)) # only the first row, which does not contain data

## Averaged over all 12 months and scaled
NPGO_full <- NPGOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(npgo = as.numeric(NPGO)) %>%
  group_by(Year) %>%
  summarise(NPGO = mean(npgo, na.rm = TRUE)) %>%
  dplyr::select(Year, NPGO) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(NPGO_full = scale(NPGO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, NPGO_full)

## Averaged for January-June only and scaled
NPGO_half <- NPGOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(npgo = as.numeric(NPGO)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>% 
  group_by(Year) %>%
  summarise(NPGO = mean(npgo, na.rm = TRUE)) %>%
  dplyr::select(Year, NPGO) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(NPGO_half = scale(NPGO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, NPGO_half)

## Averaged for December-March only and scaled
NPGO_delay <- NPGOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(npgo = as.numeric(NPGO)) %>%
  filter(Month == 12 | Month == 1 | Month == 2 | Month == 3) %>%
  filter(Year >= 2008 & Year <= 2017)%>%
  slice(-c(1:3)) %>%
  slice(-37) %>%
  mutate(Period = c(rep("A", 4), rep("B", 4), rep("C", 4), rep("D", 4), 
                     rep("E", 4), rep("F", 4), rep("G", 4), rep("H", 4), rep("I", 4))) %>%
  group_by(Period) %>%
  summarise(NPGO = mean(npgo, na.rm = TRUE)) %>%
  mutate(Year = 2009:2017) %>%
  mutate(NPGO_delay = scale(NPGO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, NPGO_delay)

## join 2 datasets together
NPGO1 <- full_join(NPGO_full, NPGO_half, by = "Year")
NPGO2 <- full_join(NPGO1, NPGO_delay, by = "Year")

## join with egg size data
NPGOdf1 <- left_join(egg, NPGO2, by = "Year")

## run models
NPGO_full_mod <- lmer(Size ~ NPGO_full + (1 | Plot), data = NPGOdf1) 
NPGO_half_mod <- lmer(Size ~ NPGO_half + (1 | Plot), data = NPGOdf1)
NPGO_delay_mod <- lmer(Size ~ NPGO_delay + (1 | Plot), data = NPGOdf1)

## compare AIC
NPGO_full_AIC <- AIC(NPGO_full_mod)
NPGO_half_AIC <- AIC(NPGO_half_mod)
NPGO_delay_AIC <- AIC(NPGO_delay_mod)

## True or False, is the full NPGO model AIC smaller than the half NPGO model AIC?
NPGO_full_AIC < NPGO_half_AIC

## What is the delta AIC?
NPGO_half_AIC - NPGO_full_AIC

## True or False, is the delay NPGO model AIC smaller than the full NPGO model AIC?
NPGO_delay_AIC < NPGO_full_AIC

## What is the delta AIC?
NPGO_delay_AIC - NPGO_full_AIC
```
The model that used all data from 12 months averaged for NPGO had the best fit. 

Then, using the full NPGO model, we tested a non-lagged versus one-year lagged dataset.
```{r NPGO2}
## no lag, data averaged for 12 months (same as NPGO_full above) and scaled
NPGO_nolag <- NPGOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(npgo = as.numeric(NPGO)) %>%
  group_by(Year) %>%
  summarise(NPGO = mean(npgo, na.rm = TRUE)) %>%
  dplyr::select(Year, NPGO) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(NPGO_nolag = scale(NPGO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, NPGO_nolag)

## one year lag, data averaged for 12 months and scaled
NPGO_lag <- NPGOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(npgo = as.numeric(NPGO)) %>%
  group_by(Year) %>%
  summarise(NPGO = mean(npgo, na.rm = TRUE)) %>%
  dplyr::select(Year, NPGO) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(NPGO_lag = scale(NPGO),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, NPGO_lag)

## join 2 datasets together
NPGO2 <- full_join(NPGO_nolag, NPGO_lag, by = "Year")

## join with egg size data
NPGOdf2 <- left_join(egg, NPGO2, by = "Year")
  
## run models
NPGO_nolag_mod <- lmer(Size ~ NPGO_nolag + (1 | Plot), data = NPGOdf2) 
NPGO_lag_mod <- lmer(Size ~ NPGO_lag + (1 | Plot), data = NPGOdf2)

## compare AIC
NPGO_nolag_AIC <- AIC(NPGO_nolag_mod)
NPGO_lag_AIC <- AIC(NPGO_lag_mod)

## True or False, is the no lag NPGO model AIC smaller than the NPGO lag AIC?
NPGO_nolag_AIC < NPGO_lag_AIC

## What is the delta AIC?
NPGO_lag_AIC - NPGO_nolag_AIC
```
The non-lagged NPGO data has more support than the one-year lagged data. There is the most support for a non-lagged NPGO data from the entire year (12 months). 

## PDO
First, we tested against using data from the entire year versus from January-June. 
```{r PDO1}
PDOraw <- read_csv(here("data", "covariates", "cciea_OC_PDO_712b_5843_9069.csv"))

## are there NAs in the data?
sum(is.na(PDOraw$PDO)) # only the first row, which does not contain data

## data averaged across entire year and scaled
PDO_full <- PDOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(pdo = as.numeric(PDO)) %>%
  group_by(Year) %>%
  summarise(PDO = mean(pdo, na.rm = TRUE)) %>%
  dplyr::select(Year, PDO) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(PDO_full = scale(PDO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, PDO_full)

## data averaged for January-June only and scaled
PDO_half <- PDOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(pdo = as.numeric(PDO)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>%
  group_by(Year) %>%
  summarise(PDO = mean(pdo, na.rm = TRUE)) %>%
  dplyr::select(Year, PDO) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(PDO_half= scale(PDO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, PDO_half)

## join 2 datasets together
PDO1 <- full_join(PDO_full, PDO_half, by = "Year")

## join with egg size data
PDOdf1 <- left_join(egg, PDO1, by = "Year")

## run models
PDO_full_mod <- lmer(Size ~ PDO_full + (1 | Plot), data = PDOdf1) 
PDO_half_mod <- lmer(Size ~ PDO_half + (1 | Plot), data = PDOdf1)

## compare AIC
PDO_full_AIC <- AIC(PDO_full_mod)
PDO_half_AIC <- AIC(PDO_half_mod)

## True or False, is the Jan-Jun PDO model AIC smaller than the PDO full AIC?
PDO_half_AIC < PDO_full_AIC

## What is the delta AIC?
PDO_half_AIC - PDO_full_AIC
```
The full PDO model has more support than the half model. 

Then, using the full PDO model, we tested a non-lagged versus one-year lagged dataset.
```{r PDO2}
## no lag, averaged for 12 months (same as PDO_full above) and scaled
PDO_nolag <- PDOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(pdo = as.numeric(PDO)) %>%
  group_by(Year) %>%
  summarise(PDO = mean(pdo, na.rm = TRUE)) %>%
  dplyr::select(Year, PDO) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(PDO_nolag = scale(PDO)) %>%
  arrange(Year) %>%
  dplyr::select(Year, PDO_nolag)

## one year lag, averaged for 12 months and scaled
PDO_lag <- PDOraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(pdo = as.numeric(PDO)) %>%
  group_by(Year) %>%
  summarise(PDO = mean(pdo, na.rm = TRUE)) %>%
  dplyr::select(Year, PDO) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(PDO_lag = scale(PDO),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, PDO_lag)

## join 2 datasets together
PDO2 <- full_join(PDO_nolag, PDO_lag, by = "Year")

## join with egg size data
PDOdf2 <- left_join(egg, PDO2, by = "Year")
  
## run models
PDO_nolag_mod <- lmer(Size ~ PDO_nolag + (1 | Plot), data = PDOdf2) 
PDO_lag_mod <- lmer(Size ~ PDO_lag + (1 | Plot), data = PDOdf2)

## compare AIC
PDO_nolag_AIC <- AIC(PDO_nolag_mod)
PDO_lag_AIC <- AIC(PDO_lag_mod)

## True or False, is the no lag PDO model AIC smaller than the PDO lag AIC?
PDO_nolag_AIC < PDO_lag_AIC

## What is the delta AIC?
PDO_nolag_AIC - PDO_lag_AIC
```
There is support for including a one-year lag for the PDO data averaged across 12 months.

## ONI
First, we tested against using data from the entire year versus from January-June. 
```{r ONI1}
ONIraw <- read_csv(here("data", "covariates", "cciea_OC_ONI_712b_5843_9069.csv"))

## are there NAs in the data?
sum(is.na(ONIraw$ONI)) # only the first row, which does not contain data

## data averaged across 12 months and scaled
ONI_full <- ONIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time)) %>%
  mutate(oni = as.numeric(ONI)) %>%
  group_by(Year) %>%
  summarise(ONI = mean(oni, na.rm = TRUE)) %>%
  dplyr::select(Year, ONI) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(ONI_full = scale(ONI)) %>%
  arrange(Year) %>%
  dplyr::select(Year, ONI_full)

## data averaged for January-June only and scaled
ONI_half <- ONIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(oni = as.numeric(ONI)) %>%
  filter(Month == 1 | Month == 2 | Month == 3 |Month == 4 | Month == 5 | Month == 6) %>%
  group_by(Year) %>%
  summarise(ONI = mean(oni, na.rm = TRUE)) %>%
  dplyr::select(Year, ONI) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(ONI_half = scale(ONI)) %>%
  arrange(Year) %>%
  dplyr::select(Year, ONI_half)

## join 2 datasets together
ONI1 <- full_join(ONI_full, ONI_half, by = "Year")

## join with egg size data
ONIdf1 <- left_join(egg, ONI1, by = "Year")

## run models
ONI_full_mod <- lmer(Size ~ ONI_full + (1 | Plot), data = ONIdf1) 
ONI_half_mod <- lmer(Size ~ ONI_half + (1 | Plot), data = ONIdf1)

## compare AIC
ONI_full_AIC <- AIC(ONI_full_mod)
ONI_half_AIC <- AIC(ONI_half_mod)

## True or False, is the Jan-Jun ONI model AIC smaller than the ONI full AIC?
ONI_half_AIC < ONI_full_AIC

## What is the delta AIC?
ONI_half_AIC - ONI_full_AIC
```
The full ONI model has more support than the half-year ONI model. 

Using the full ONI model, we tested a non-lagged versus one-year lagged dataset.
```{r ONI2}
## no lag, averaged for 12 months (same as ONI_full above) and scaled
ONI_nolag <- ONIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(oni = as.numeric(ONI)) %>%
  group_by(Year) %>%
  summarise(ONI = mean(oni, na.rm = TRUE)) %>%
  dplyr::select(Year, ONI) %>%
  filter(Year >= 2009 & Year <= 2017) %>%
  mutate(ONI_nolag = scale(ONI)) %>%
  arrange(Year) %>%
  dplyr::select(Year, ONI_nolag)

## one year lag, averaged for 12 months and scaled
ONI_lag <- ONIraw %>%
  slice(-1) %>%
  mutate(time = ymd_hms(time)) %>%
  mutate(Year = year(time),
         Month = month(time)) %>%
  mutate(oni = as.numeric(ONI)) %>%
  group_by(Year) %>%
  summarise(ONI = mean(oni, na.rm = TRUE)) %>%
  dplyr::select(Year, ONI) %>%
  filter(Year >= 2008 & Year <= 2016) %>%
  mutate(ONI_lag = scale(ONI),
         TrueYear = Year,
         Year = TrueYear + 1) %>%
  arrange(Year) %>%
  dplyr::select(Year, ONI_lag) 

## join 2 datasets together
ONI2 <- full_join(ONI_nolag, ONI_lag, by = "Year")

## join with egg size data
ONIdf2 <- left_join(egg, ONI2, by = "Year")
  
## run models
ONI_nolag_mod <- lmer(Size ~ ONI_nolag + (1 | Plot), data = ONIdf2) 
ONI_lag_mod <- lmer(Size ~ ONI_lag + (1 | Plot), data = ONIdf2)

## compare AIC
ONI_nolag_AIC <- AIC(ONI_nolag_mod)
ONI_lag_AIC <- AIC(ONI_lag_mod)

## True or False, is the no lag ONI model AIC smaller than the ONI lag AIC?
ONI_nolag_AIC < ONI_lag_AIC

## What is the delta AIC?
ONI_lag_AIC - ONI_nolag_AIC
```
The no lag full ONI model has more support than the one-year lagged full ONI model. 
